---
title: "HMM-based Futures Price Forecasts"
author: "Chris D"
date: "5/29/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Introduction
> Because my R studio version cannot complie Chinese character, I can only comment by using English. 

## Packages Preparation

- `depmixS4`: a good package for HMM model, but not as good as packages in python, API is very hard to read : https://cran.r-project.org/web/packages/depmixS4/depmixS4.pdf 

- `readxl`: a common package to read excel files

```{r warning=FALSE, echo = T, results = 'hide', message=FALSE}
library(depmixS4)
library(readxl)
```


## Load and Clean Data  

- Clean our data, I divide the file you sent to me into individual file with one specific future index

```{r setup, include=FALSE}
corn_data <- read_excel("corn_data.xls")
rubber_data <- read_excel("rubber_data.xls")
sugar_data <- read_excel("sugar_data.xls")

dataClean <- function(inputData){
  inputData <- inputData[, 3:ncol(inputData)]
  names(inputData) <- c('date', 'open', 'close', 'high', 'low')
  inputData <- data.frame(inputData)
  
  # need to specify date column
  inputData$date <-as.Date(inputData$date, "%Y-%m-%d")
  
  # omit NA rows in each dataset
  outputData <- na.omit(inputData)
  return(outputData)
}

corn_data <- dataClean(corn_data)
rubber_data <- dataClean(rubber_data)
sugar_data <- dataClean(sugar_data)
```

## Construct Model (take cornData as an example)

Outcome of 
```{r}
## Construct and fit a 4-state model
mod <- depmix(list(open ~ 1, close ~ 1, high ~ 1, low ~ 1), data=corn_data, nstates=4, family=list(gaussian(), gaussian(), gaussian(), gaussian()))

set.seed(1)
fm2 <- fit(mod, em = em.control(maxit = 1000),verbose = FALSE)

summary(fm2)
print(fm2)

probs <- posterior(fm2)             # Compute probability of being in each state
head(probs)

```

## Function Definition and Explanation

- Define a `fitHMM` function to fit assigned dataset and number of hiddenState

```{r}
fitHMM <- function(dataset, hiddenState){
  for (i in 2:hiddenState){
    set.seed(1)
    cat("Outcome of ", deparse(substitute(dataset)), "\n")
    fmx <- fit(depmix(list(open ~ 1, close ~ 1, high ~ 1, low ~ 1), data = dataset, nstates = i,
                      family=list(gaussian(), gaussian(), gaussian(), gaussian())))
    
    summary(fmx)
    print(fmx)
    poster <- posterior(fmx)
  }
}

fitHMM(sugar_data, 4)

```


